Technology for application family creation based on domain analysis

### 2016-04-09 2016-domainAnalysis/domainAnalysis.tex

\section{Domain analysis approaches}
\label{chapter:approaches}
There is no any clear and long-standing definition of the term ``domain analysis''. Almost all papers, in which this term is considered, go back to 80s-90s of the twentieth century. It was then that scientists, taking into account rapidly growing technologies, were thinking about global reuse. Always projects are developing for concrete user needs, so then the term ``domain'' took the definition. Domain is the field of expertise, problems in which the software intend to solve. According to Rugaber [1], the domain is described in terms of glossary, some assumptions, architecture approach and literature.

Then the question arise, how we need to analyze the domain for acquiring the necessary information. At present, the information gathering into knowledge bases is understood under the term ``domain analysis''. Although, Prieto-Diaz [2] confirms that domain analysis is an activity, which take place before system analysis and its output is used for system analysis to the same degree as system analysis’s output is used for system design.  There are other definitions of the term ``domain analysis''. Ferre [3] has presented definitions, such as: (1) the process of identification, organization and presenting the relevant information of a given domain, (2) the process, in which the customer’s knowledge are identified, concretized and systemized. The relevant information of the domain should be presented in objective, readily available way, such way is called ``domain model''. Mernik [4] specifies, that the domain model include not only glossary, but also must describe commonalities and variabilities of terms. Such model should precisely set bounds of the domain, i.e. clear and exact characterize a range of questions, which are considered in the domain. Term variabilities allow to define exactly, what information must be specified in concrete system implementation. Term commonalities are used for defining a set of shared operations between different applications. Implementing commonalities and adding the gained model  with information, which can be specified in instance of the concrete system, a set of different system can be obtained based only one common model. In such manner, based only one domain model can be implemented the set of different systems in given domain. Taking into account definitions above mentioned, we can conclude that domain analysis is the activity forward system analysis, which goal is providing the domain model.

@@

Next step is to create domain-specific language that allows to combine and configure features from feature library to implement applications in given domain. This is where our approach differs from common reuse strategies. Naive approach would be to generate an application directly from feature diagram, somehow marking features that shall be included into application, and it actually works fine when domain variability is low. But more common is the situation when features themselves have properties that allow to configure them, those properties can have different types. Also, components may be related to each other in different ways, be used in configuration of one another, or some of their properties may be meaningless in absence or presence of other feature. Those rules may be implemented implicitly in application generator and require that programmers will always observe them, but we propose that these rules will be captured explicitly by dedicated domain-specific language. Such language may make models that do not observe those rules syntactically incorrect, and it will greatly reduce the possibility of human error and reduce knowledge required to efficiently use programming system.

@@

\section{Conclusion}
\label{chapter:conclusion}
The problem of not using domain analysis result for further generating some entities for software development process was stated.There were considered some formal domain analysis approaches and we concluded that creation of feature diagrams is the most elegant decision for domain analysis that can be conducted by domain expert, i.e. non-programmer, may be in collaboration with system analysts. Moreover, there was discussed one of the possible solutions, which is presented by Estublier, we specify some problems of such method. We suggested our own approach for creation application family in one domain based on domain analysis. Also, there was some evaluation of this approach, where we pointed out that this example remains many questions because of its simplicity. Finally, this research requires more detailed explorations.

>>>>

\section{Domain analysis approaches}
\label{chapter:approaches}
There is no any clear and long-standing definition of the term ``domain analysis''. Almost all papers, in which this term is considered, go back to 80s-90s of the twentieth century. It was then that scientists, taking into account rapidly growing technologies, were thinking about global reuse. Always projects are developing for concrete user needs, so then the term ``domain'' took the definition. Domain is the field of expertise, problems in which the software intend to solve. According to Rugaber~\cite{rugaber1994domain}, the domain is described in terms of glossary, some assumptions, architecture approach and literature.

Then the question arise, how we need to analyze the domain for acquiring the necessary information. At present, the information gathering into knowledge bases is understood under the term ``domain analysis''. Although, Prieto-Diaz~\cite{prieto1988domain} confirms that domain analysis is an activity, which take place before system analysis and its output is used for system analysis to the same degree as system analysis’s output is used for system design.  There are other definitions of the term ``domain analysis''. Ferre~\cite{ferre1999evaluation} has presented definitions, such as: (1) the process of identification, organization and presenting the relevant information of a given domain, (2) the process, in which the customer’s knowledge are identified, concretized and systemized. The relevant information of the domain should be presented in objective, readily available way, such way is called ``domain model''. Mernik~\cite{mernik2005} specifies, that the domain model include not only glossary, but also must describe commonalities and variabilities of terms. Such model should precisely set bounds of the domain, i.e. clear and exact characterize a range of questions, which are considered in the domain. Term variabilities allow to define exactly, what information must be specified in concrete system implementation. Term commonalities are used for defining a set of shared operations between different applications. Implementing commonalities and adding the gained model  with information, which can be specified in instance of the concrete system, a set of different system can be obtained based only one common model. In such manner, based only one domain model can be implemented the set of different systems in given domain. Taking into account definitions above mentioned, we can conclude that domain analysis is the activity forward system analysis, which goal is providing the domain model.

@@

Next step is to create domain-specific language that allows to combine and configure features from feature library to implement applications in given domain. This is where our approach differs from common reuse strategies. Naive approach would be to generate an application directly from feature diagram, somehow marking features that shall be included into application, and it actually works fine when domain variability is low~\cite{she2010variability}. But more common is the situation when features themselves have properties that allow to configure them, those properties can have different types. Also, components may be related to each other in different ways, be used in configuration of one another, or some of their properties may be meaningless in absence or presence of other feature. Those rules may be implemented implicitly in application generator and require that programmers will always observe them, but we propose that these rules will be captured explicitly by dedicated domain-specific language. Such language may make models that do not observe those rules syntactically incorrect, and it will greatly reduce the possibility of human error and reduce knowledge required to efficiently use programming system.

@@

\section{Conclusion}
\label{chapter:conclusion}
The problem of not using domain analysis result for further generating some entities for software development process was stated. There were considered some formal domain analysis approaches and we concluded that creation of feature diagrams is the most elegant decision for domain analysis that can be conducted by domain expert, i.e. non-programmer, may be in collaboration with system analysts. Moreover, there was discussed one of the possible solutions, which is presented by Estublier, we specify some problems of such method. We suggested our own approach for creation application family in one domain based on domain analysis. Also, there was some evaluation of this approach, where we pointed out that this example remains many questions because of its simplicity. Finally, this research requires more detailed explorations.
